---
title: "API"
author: "Olga Shiligin"
date: "27/10/2018"
output: html_document
---

The New York Times web site provides a rich set of APIs, as described here: http://developer.nytimes.com/docs You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

I have selected article search API with search query - Brexit as it is very "hot" topic in the UK now.

Loading necessary library

```{r}
library(jsonlite)
```

Signing up for an API key

```{r}
api_key<-"?api-key=ec35a8674ec64bc2b24217d98784448a"
```

Setting up search criterias

```{r}
q<-"&q=brexit"
begin_date<-"&begin_date=20180901"
end_date<-"&end_date=20181020"
page<-"&page="
```

Reading JSON and transforming data to R data frame

```{r}
df<-data.frame()
 # getting 30 page (10 results per page)
for (i in 1:30){
  # constructing URL from search criterias
    url <- paste0('https://api.nytimes.com/svc/search/v2/articlesearch.json',api_key,q,begin_date,end_date,page,toString(i))
    data<-fromJSON(url,flatten = TRUE)
    # selecting only necessary columns from JSON file
    docs.df <- data$response$docs
    for(k in 1:nrow(docs.df)){
      temp_df <- data.frame(headline=docs.df[k,"headline.main"],source=docs.df[k,"source"],section=docs.df[k,"section_name"],news_description=docs.df[k,"news_desk"],document_type=docs.df[k,"document_type"])
      df <- rbind(df,temp_df)
    }
    Sys.sleep(2)
    print(paste("finished page ",i))
}
head(df)
```

