---
title: "Movies"
author: "Olga Shiligin"
date: "07/09/2018"
output: html_document
---

Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.
Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.


Installing RMySQL, DBI, connecting to MySql Database

```{r}
# install.packages("RMySQL")
# install.packages("DBI")
library(RMySQL)

conbection <- dbConnect(MySQL(),
                 user="root", password="olga123",
                 dbname="607_assignments", host="localhost")
```


```{r}
results <- "SELECT * from movies"
results <- dbGetQuery(conbection, results)
results
```

Selecting movie with the highest rating

```{r}
best_movie <- "SELECT Movie, avg(Rating) from movies group by Movie order by avg(Rating) desc"
best_movie <- dbGetQuery(conbection, best_movie)
best_movie

```
Selecting person who loves the selected movies most.

```{r}
person <- "SELECT Name, avg(Rating) from movies group by Name order by avg(Rating) desc"
person <- dbGetQuery(conbection, person)
person

```
Creating R dataframe.

```{r}

df<-data.frame(results)
head(df)

```

Disconnecting from database.

```{r}
dbDisconnect(conbection)

```











