---
title: "607_Project1"
author: "Olga Shiligin"
date: "23/09/2018"
output: html_document
---

Task
----
The purpose of the project is to process a text file of chess tournament results and generate a csv file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

Solution
----------

"stringr" package is required for the project

Reading table using read.table function. One extra column (empty) was created because "|" is treated as the beginning of the next column. We can delete the empty column.

```{r}
library(stringr)
column_names <- c("ID", "Player_Name", "Total_Ptc", "R1", "R2", "R3", "R4", "R5", "R6", "R7", "Empty")
chess_results <- read.table("https://raw.githubusercontent.com/olgashiligin/Chess-Tournament/master/Raw%20Data", header = FALSE, sep = "|", col.names = column_names, skip = 4, fill = TRUE, stringsAsFactors = FALSE)
chess_results <- subset(chess_results,select = c(1:10))
head(chess_results,5)
```

```{r}
is.data.frame(chess_results)
```

Now we have to remove empty spaces across all the table.

```{r}
for (row in 1:nrow(chess_results)) {
  for(col in 1:10) {
    chess_results[row,col] <- str_trim(chess_results[row,col])
  }
}
```

Creating State and Pre_Score columns

```{r}
for (row in 1:nrow(chess_results)) {
  chess_results$State[row] <- chess_results$ID[row+1]
  chess_results$Pre_Score[row] <- str_extract(chess_results$Player_Name[row+1], "[:space:]{1}\\d{3,4}")
}

chess_results <- subset(chess_results, !is.na(chess_results$Pre_Score))

head(chess_results,5)

```

The following steps to be taken in order to calculate opponent's average pre rating.

1. Removing letters (W, L, D) from R1 to R7 columns (we do not need these letters in our project)
2. Replacing opponent's ID in Rs columns with opponent's pre rating.
3. Calculating opponent's average pre rating.

Removing letters:

```{r}
for (col in 4:10) {
  chess_results[, col] <- str_extract(chess_results[, col], "\\d+")
}

head(chess_results,5)
```

Replacing opponent's ID in Rs columns with opponent's pre rating.

```{r}
for (row in 1:nrow(chess_results)) {
  for(col in 4:10) {
    filter<-chess_results$ID == chess_results[row,col]
    chess_results[row,col] <- chess_results[filter,12][1]
  }
}

head(chess_results,5)

```

Calculating opponent's average pre rating.
Checking data frame structure first.

```{r}
str(chess_results)
```

Changing data type for Rounds in order to calculate Average pre rating for opponets.

```{r}
for (col in 4:10) {
  chess_results[, col] <- as.numeric(chess_results[, col])
}
```

Ceating new column Opponents_Rating
https://stackoverflow.com/questions/33981527/compute-the-mean-of-two-columns-in-a-dataframe

```{r}
chess_results$Opponent <- round(rowMeans(chess_results[, c(4:10)], na.rm = TRUE), 0)
```

Final Data Frame

```{r}
chess_results <- subset(chess_results,select = c(1,2,3,11,12,13))
head(chess_results,5)
```

Saving Results to CSV file

```{r}
write.csv(chess_results, "chess_results.csv", row.names=FALSE)
```
