---
title: "RegEx"
author: "Olga Shiligin"
date: "16 September 2018"
output: html_document
---

```{r}
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"
names<- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
names
```
(a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.
 2 steps:
 - remove titles/middle name;
 - change first/last name position with "," inside.
 
```{r}
title_free<- str_trim(sub("[[:alpha:]]{1,3}\\.","",names)) 
first_last<- str_replace(title_free,"(\\w+),\\s+(\\w+)","\\2 \\1")
first_last
```

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.)

```{r}
has_title <- str_detect(names, "^[:alpha:]{2,3}\\.") 
title<-data.frame (names, title=has_title)
title
```

(c) Construct a logical vector indicating whether a character has a second name.

```{r}
second_name <- str_detect(names, " [:alpha:]\\.") 
second_name<-data.frame (names, second_name=second_name)
second_name
```

4. Describe the types of strings that conform to the following regular expressions and
construct an example that is matched by the regular expression.

(a) [0-9]+\\$ 
Any numbers or span of numbers from 0 to 9 followed by a $

```{r}
str_detect ("839503$", "[0-9]+\\$")
str_detect ("3$", "[0-9]+\\$")
```

(b) \\b[a-z]{1,4}\\b
Words that are constructed from 1 to 4 letters (lower case) surrounded by the word edge.

```{r}
str_detect("buzz","\\b[a-z]{1,4}\\b")
str_detect("z","\\b[a-z]{1,4}\\b")
```

(c) .*?\\.txt$
It can be anything that ends with .txt

```{r}
str_detect(".txt",".*?\\.txt$")
str_detect("jslfn375hfv .txt",".*?\\.txt$")
str_detect(" .txt",".*?\\.txt$")
```

d) \\d{2}/\\d{2}/\\d{4}
Date format dd/mm/yyyy

```{r}
str_detect("13/09/2018","\\d{2}/\\d{2}/\\d{4}")
```

(e) <(.+?)>.+?</\\1>
 HTML tag
 
```{r}
 str_detect("<<td>Defines the document type</td>>","<(.+?)>.+?</\\1>")
```
 
9. The following code hides a secret message.

```{r}
secret<- ("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6.Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
secret <- unlist(str_extract_all(secret, "[[:upper:].?]{1,}"))
message <- str_replace_all(paste(secret, collapse = ''), "[.]", " ")
message
```

