---
title: "Project2_Athletes"
author: "Olga Shiligin"
date: "07/10/2018"
output: html_document
---

Data Source: https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results

Data Set:“120 years of Olympic history: athletes and results”. Basic bio data on athletes and medal results from Athens 1896 to Rio 2016.

1. Installing necessary packages

```{r}
# install.packages("tidyr")
# install.packages("dplyr")
# install.packages("ggplot2")
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)
```

2. Reading csv file (file is too big for placing it on GitHub)

```{r}
athletes<- read.csv("/Users/Olga/Desktop/athlete_events.csv")
```

3. Subsetting dataset: selecting only necessary for the analysis columns, filtering rows which do not have medal type.

```{r}
sub_athletes<-athletes %>% 
  select(-c(NOC,Games,Event)) %>% 
   filter (Medal != "<NA>") 
head(sub_athletes)
```

Genaral Analysis

a) What counties have the biggest number of medals over the whole olympic history?

```{r}
total_winners<-sub_athletes %>% 
  filter (Medal != "<NA>") %>% 
   group_by(Team) %>% 
   count(Medal) %>% 
  mutate(sum_n = sum(n)) %>% 
  select(Team,sum_n) %>% 
  slice(1) %>% 
  arrange(desc(sum_n)) 

total_winners 

```

b) What kind of sport and what countries have the largest number of gold medals over the whole olympic history?

```{r}
sport_winners<-sub_athletes %>% 
  filter(Medal=="Gold") %>% 
  group_by(Sport,Team) %>% 
  count(Medal) %>%
  arrange(desc(n))

head(sport_winners)
tail(sport_winners)

```


c) What country has the highest number of gold medals by each Olympic games.

```{r}
winner_by_year<-sub_athletes %>%
  filter(Medal =="Gold") %>%
  select(Year,Team,Medal) %>% 
  group_by(Year,Team) %>% 
  count(Medal) %>%
  group_by(Year) %>% 
  arrange(Year,desc(n)) %>% 
  slice(1)
 
head(winner_by_year)

```

d) What country appear most often across all Olympic games history as a gold medal winner?

```{r}
absolute_leader<-winner_by_year %>% 
  group_by(Team) %>% 
  count(Team) %>% 
  arrange(desc(nn))
  
absolute_leader
```

e) Any correlation between kind of sports and height?

```{r}
height_sport<-sub_athletes %>% 
  filter (Height != "<NA>") %>% 
  group_by(Sport) %>% 
  summarise(avg_height=mean(Height)) %>% 
  arrange(desc(avg_height))

head(height_sport)
tail(height_sport)

```

f) Any correlation between kind of sports and weight?

```{r}
Weight_sport<-sub_athletes %>% 
  filter (Weight != "<NA>") %>% 
  group_by(Sport) %>% 
  summarise(avg_Weight=mean(Weight)) %>% 
  arrange(desc(avg_Weight))

head(Weight_sport)
tail(Weight_sport)

```

