---
title: "Tidying and Transforming Data"
author: "Olga Shiligin"
date: "29/09/2018"
output: html_document
---

Installing necessary packages
```{r}
# install.packages("tidyr")
# install.packages("dplyr")
# install.packages("ggplot2")
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)

```

Reading data from csv file:
---------------------------
```{r}
data<- read.csv("https://raw.githubusercontent.com/olgashiligin/Tidying_and_Transforming/master/delays.csv",sep=",")
data

```
Tidying Data:
---------------
1. Removing one empty row
2. Renaming first two columns
3. Replacing empty spaces with Airlines names


 Removing one empty row

```{r}
data<-data[-c(3), ] 
data
```


  Renaming first two columns

```{r}
names(data) <- c('Airlines', 'Status', 'Los Angeles', 'Phoenix', 'San Diego','San Francisco', names(data)[7])
data
```

 Replacing empty spaces with Airlines names

```{r}
for (i in 1:nrow(data)) {
  if (data[i, 1] == " ") {
    data[i, 1] = data[i - 1, 1]
  }
}

data

```
Prepairing data set for the analysis:
--------------------------------------
1. Transforming wide table format
2. Splitting "Status" column into two columns: delayed and on_time.


 Transforming wide table format

```{r}
data<-gather(data,Destinations,Number_of_Flights,3:7)
data
```

 Splitting "Status" column into two columns: delayed and on_time. Renaming on time column.

```{r}
data <- spread(data, Status, Number_of_Flights)
names(data)[names(data) == "on time"] <- "on_time"
data
```

 Performing analysis on flight delays by ALASKA and AM WEST airlines. 
---------------------------------------------------------------------
 Calculating percentage of flight delays by airlines. 

```{r}
overall_analysis<-data %>% 
   mutate(total_flights = delayed + on_time) %>%
    group_by(Airlines) %>% 
     summarise(avg_delayed_rate = sum(delayed)/sum(total_flights))

overall_analysis

```
Overall analysis shows that Alaska airlines has higer rate of delays (0.133) compare to  AM WEST (0.109)

```{r}
analysis<-data %>%
  mutate(total_flights = delayed + on_time, delayed_rate = delayed/(delayed+on_time)) %>%
  arrange(desc(delayed_rate))

analysis
```


 
```{r}
viz3<-ggplot(analysis, aes(x = Destinations, y = delayed_rate, fill = Airlines)) + geom_col( position = "dodge") + ylab("Delayed Flights, %")+ggtitle("           Flight Delays By Airlines and By Destinations")
viz3

```

AM WEST: the highest rate of delays - San Francisco (0.28), the lowest rate of delays - Phoenix (0.79)

Alaska: the highest rate of delays - San Francisco (0.17), the lowest rate of delays - Phoenix (0.05)






