---
title: "Final Project"
author: "Olga Shiligin"
date: "08/12/2018"
output: html_document
---

Innovations Of Top 10 Technology Companies
-------------------------------------------

Introduction
-------------

We all know that patents is one of the forces driving progress, I was curious how leading tech companies perform in this field. Do they innovate a lot? How innovations rate fluctuates? What is the most popular category of innovation? What are global innovation trends?


Goal
----
For my final project I will perform a study analysis of the patent registration data of the top 10 Technology Companies in the world trying to find answers to the following questions:

- what is the overall innovation trend of top 10 tech companies?
- how many patents tech companies register during their entire history, who is at the top of this list?
- how do companies' main innovation categories change over time?
- what are the area of main focus for innovation of each company?


Data Sources
------------

1. Patents registration data is taken from US Patents And Trademark Office website: https://www.uspto.gov/patents-application-process/search-patents
2. Patents US class classification: https://www.uspto.gov/web/patents/classification/selectnumwithtitle.htm
3. Top 10 Top 10 Technology Companies (2018): https://www.investopedia.com/articles/markets/030816/worlds-top-10-technology-companies-aapl-googl.asp


Limitations and assumptions:
----------------------------

1. The defenition of a ‘digital tech business’ is a business that provides a digital technical service/product/platform/hardware, or heavily relies on it, as its primary revenue source. 
As the majority of medium to large companies use and heavily rely on technology. I would consider company as a Technology Company if it is in the business of SELLING technology. For the purpose of the project I took the list top 10 Technology Companies by Investopedia list.

https://www.investopedia.com/articles/markets/030816/worlds-top-10-technology-companies-aapl-googl.asp

- Apple (Market value: $995.50 billion)
- Samsung (Market value: $765.26 billion)
- Microsoft (Market value: $855.4 billion)
- Google (Alphabet) (Market value: $756.85 billion)
- Intel (Market value: $220.33 billion)
- IBM (Market value: $112.47 billion)
- Facebook (Market value: $434.66 billion)
- Hon Hai Precision (Market cap: $6023.96 billion)
- Tencent (Market value: $356.11 billion)
- Oracle (Market value: $190.38 billion)

2. IMPORTANT: As patent registrtration process takes up to 2 years to be fully registred in US Patents And Trademark Office, The data for the last 2 years is ignored in my analysis.


Technical project overview
---------------------------

1. Scraping patent registration data for the whole company history. Python language was used for data scraping.
 
 Example of web page where data was scraped:
 
http://patft.uspto.gov/netacgi/nph-Parser?Sect1=PTO2&Sect2=HITOFF&p=1&u=%2Fnetahtml%2FPTO%2Fsearch-bool.html&r=37&f=G&l=50&co1=AND&d=PTXT&s1=Apple.ASNM.&OS=AN/Apple&RS=AN/Apple

The following data was scraped for each company:

- United States Patent Number

- Date Field

- Current U.S. Class

- Patent title

- Assignee name

- link

2. Data Transformation: change initial data structure to data structure that is suitable and ready to use in relational database. Python language was used to change data structure.

 - current U.S. Class was splitted into 2 columns (before and after slash): us_category and us_subcategory. Only us_category was used in the analysis.
 - year was extracted from data column
 - companies name was added ("author" column)
 - only top 4 columns of scraped data will be used for my analysis (patent_number, date, us_cat, title)
 
3. Saving data to csv files. Python language was used to save data in csv files.

4. Importing csv file to MySql database. R language was used for the importing.

5. Reading data from MySql database. R language and SQL language was used for reading data from database.

6. Performing analysis. (R language)


Data Collection Statistics Overview:
------------------------------------

Time period: 1968 till present

Number of patents regestred: 353683

Number of categories: 436

Number of companies analysed: 10


Note: Data Collection and Transformation Script is written in Python and presented in GitHub. The rest script is presented below.



```{r}
library("stringr")
library("dplyr")
library("tidyr")
library("RMySQL")
library(DBI)
library("ggplot2")
#install.packages("sqldf")
#library("sqldf")
library(plyr)
library(wordcloud)

#declare some constants
data.dir = "/Users/Olga/Desktop/PycharmProjects/r/data/"
input.files.dir = paste0(data.dir,"patents/")
classification.path = paste0(data.dir,"classification_csv.csv")
suffix = "_us_csv.csv"

#function for connecting to MySQL database
create.connection <- function(){
  connection <- dbConnect(MySQL(),
                          user="root", password="olga1234",
                          dbname="patents", host="localhost")
}

#function for creating table in db
create.table <- function(connection, schema, table){
  create.table.query = paste0("CREATE TABLE ",schema,".",table,"
                              (
                              author        varchar(100)  not null,
                              patent_number varchar(20)   not null,
                              date          varchar(20)   null,
                              year          varchar(20)   null,
                              us_cat        varchar(200)  null,
                              us_subcat     varchar(200)  null,
                              description   varchar(1000) null,
                              title         varchar(1000) null
                              )"
                         )
  created.table <- dbSendStatement(connection, create.table.query)
  dbClearResult(created.table)
}
#function for dropping db table
drop.table <- function(connection, schema, table){
  drop.table.query = paste0("DROP TABLE IF EXISTS ",schema,".",table)
  drop.table <- dbSendStatement(connection, drop.table.query)
  dbClearResult(drop.table)
}

read.db <- function(table){
  connection <- create.connection()
  query <- paste0("SELECT * FROM ",table)
  results <- dbGetQuery(connection,query)
  results
}
#function for escaping single quote
clense <- function(val){
  gsub("'", "''",toString(val))
}

#function for saving multiple records to db in a single transactopn
save.to.db <- function(connection,patents.df,author){
  for (row in 1:nrow(patents.df)) {
    patent.number <- clense(patents.df[row, "patent_number"])
    date  <- clense(patents.df[row, "date"])
    year  <- clense(patents.df[row, "year"])
    us_cat <- clense(patents.df[row,"us_cat"])
    us_subcat <- clense(patents.df[row,"us_subcat"])
    title <- clense(patents.df[row,"title"])
    description <- clense(patents.df[row,"description"])
    insert.query = paste("INSERT INTO patents.PATENTS VALUES ( '",author,"','",patent.number,"', '",date,"', '",year,"', '",us_cat,"', '",us_subcat,"', '",description,"', '",title ,"')",sep="")
    insert.row <-  dbSendStatement(connection, insert.query)
    dbClearResult(insert.row)
  }
}
# function for reading csv
read_csv <- function(path,connection){
  tryCatch({read.csv(file=path, fileEncoding ="cp1251", header=TRUE, quote = "\"", sep=",",stringsAsFactors=FALSE)}
           , error = function(e) {
             print(paste0("following file causing error: ",path))
             print(e)
             dbDisconnect(connection)
             stop(e)
           })
}
#this function does all the work with help of all functions declared above
saveall <- function(){
  connection <- create.connection()
  pattern = paste0(suffix,"$")
  files.list <- list.files(input.files.dir,pattern = pattern)
  classification.df = read_csv(classification.path, connection)
  #colnames(classification.df)[1] <- "cat"
  for (path in files.list){
    absolute.path = paste0(input.files.dir,"/",path)
    data.df <- read_csv(absolute.path, connection)
    df <-  left_join(data.df, classification.df, by = c('us_cat'))
    author = tolower(str_replace(path, suffix,"")[[1]][1])
    tryCatch({save.to.db(connection, df,author)}
             , error = function(e) {
               print(paste0("following file causing error: ",path))
               print(e)
               dbDisconnect(connection)
               stop(e)
             }
    )
  }
}

#saveall()
```


Analysis
--------

```{r}
connection <- create.connection()
query<- "SELECT * from patents.PATENTS"
results <- dbGetQuery(connection,query)
dbDisconnect(connection)
```

General innovation pace overview by companies by year
------------------------------------------------------

```{r}
df<-results %>% 
  select (author, year, patent_number) %>% 
  group_by(author, year) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number))
```


```{r}
df$year<- as.numeric(as.character(df$year))
overall<-ggplot(data = df, aes(x=year, y = patent_count, color = author)) +
         geom_line()+ theme(axis.text.x = element_text(angle = 90, hjust =1))+scale_color_brewer(palette="Spectral")+
         ggtitle ("Number Of Registered Patents By Company By Year") + xlab("year") + ylab("number of patents")
overall
```

Noticable leaders of innovations among top 10 companies are IBM and Samsung. 

Intel and Microsoft have steady decreasing pace in patent registrations.

Facebook and Tencent innovate least of all companies.

Google has a sharp increase in patent registration since 2010 and demonstrate good innovation potential.


Aggregated innovation pace overview by year (all 10 companies)
--------------------------------------------------------------

```{r}
df1<-results %>% 
  select (year, patent_number) %>% 
  group_by(year) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number))
```


```{r}
df1$year<- as.numeric(as.character(df1$year))
overall_agg<-ggplot()+
  geom_line(data = df1, aes(x = year, y = patent_count), color = "blue") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle ("Number Of Registered Patents By Year (top 10 tech companies)") + xlab("year") + ylab("number of patents")
overall_agg
```

Graph shows that sharp increase in innovation pace starts from 1990. 

Most likely following crises affected innovation pace of top 10 tech companies:

- dot-com bubble crisis slowed the pace a bit, but after that steady growth have being continued.

- Economic crisis 2007-2008 significantly negatively affected the innovation pace and only in 2012 innovation pace exceed the pace of pre crisis pace.


What do tech companies innovate most (top 20 most popular patent categories)?
-----------------------------------------------------------------------------

```{r}
df2<-results %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=20)
```


```{r}
popular_title<-ggplot(df2, aes(x=reorder(description,-patent_count), y=patent_count)) +
  geom_bar(stat="identity")+theme_minimal() +
  geom_col(aes(fill = patent_count)) + 
  scale_fill_gradient2(low = "white", high = "blue") + 
  coord_flip() + 
  theme(text = element_text(size=6))+
  ggtitle ("The most popular patents of all time among top 10 tech companies") + xlab("Total Count") + ylab("Patent Group Name")

popular_title
```

If I ignore Miscellaneous category as it tells a litle about patent nature, the leading patent categories are active solid-atate devices; data processing: database and file management; electrical computers and digital processing systems.


Evolution Of Innovations Of Top 10 companies (from 1968 till present)
--------------------------------------------------------------------

In order to admit any changes in innovation trend over time, I will look at the leading innovation categories of the following periods:

- before 1990 year

- from 1990 till 2000

- from 2000 till 2010

- 2010 till present

Leading innovation categories before 1990 year
----------------------------------------------

```{r}
results$year<- as.numeric(as.character(results$year))
df3<-results %>% 
  filter (year<1990) %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df3)
```

Leading innovation categories from 1990 till 2000
--------------------------------------------------

```{r}
results$year<- as.numeric(as.character(results$year))
df4<-results %>% 
  filter(between(year,1990,2000)) %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df4)
```

Leading innovation categories from 2000 till 2010
--------------------------------------------------

```{r}
results$year<- as.numeric(as.character(results$year))
df5<-results %>% 
  filter(between(year, 2000 , 2010 )) %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df5)
```

Leading innovation categories from 2010 till 2018
--------------------------------------------------

```{r}
results$year<- as.numeric(as.character(results$year))
df6<-results %>% 
  filter(between(year, 2010 , 2018 )) %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df6)
```

Main innovation categories change over time, but "Active solid state devices" is the leading patent category across all periods.
Also the "evolution" of data related leading patent categories is the following: starting from data/information storage before 90th, then data processing: database and file management (1990-2020) to Recording, communication, or information retrieval equipment and image processing (from 2010 till present).


Leading innovation categories by companies
------------------------------------------

Apple
-----

```{r}
results$year<- as.numeric(as.character(results$year))
df7<-results %>% 
  filter(author =="apple") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df7)
```


Facebook
--------

```{r}
results$year<- as.numeric(as.character(results$year))
df8<-results %>% 
  filter(author =="facebook") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df8)
```

Google
-------

```{r}
results$year<- as.numeric(as.character(results$year))
df9<-results %>% 
  filter(author =="google") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df9)
```

IBM
---

```{r}
results$year<- as.numeric(as.character(results$year))
df10<-results %>% 
  filter(author =="ibm") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df10)
```

Samsung
-------

```{r}
results$year<- as.numeric(as.character(results$year))
df11<-results %>% 
  filter(author =="samsung") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df11)
```

Intel
-----

```{r}
results$year<- as.numeric(as.character(results$year))
df12<-results %>% 
  filter(author =="intel") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df12)
```

Tencent
-------

```{r}
results$year<- as.numeric(as.character(results$year))
df13<-results %>% 
  filter(author =="tencent") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df13)
```

Oracle
------

```{r}
results$year<- as.numeric(as.character(results$year))
df14<-results %>% 
  filter(author =="oracle") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df14)
```

Hon Hai Precision
-----------------

```{r}
results$year<- as.numeric(as.character(results$year))
df15<-results %>% 
  filter(author =="hon_hai") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df15)
```

Microsoft
---------

```{r}
results$year<- as.numeric(as.character(results$year))
df16<-results %>% 
  filter(author =="microsoft") %>% 
  select (description,patent_number) %>% 
  group_by(description) %>% 
  dplyr::summarise(patent_count = n_distinct(patent_number)) %>% 
  arrange(desc(patent_count)) %>% 
  top_n(n=10)
head(df16)
```

Conclusions
-----------

- Samsung, IBM and Microsoft are innovation leaders among top 10 tech companies selected for the purpose of this project.

- Intel and Microsoft have noticeable decrease in thier innovation pace, where as google demonstrate steady increasing innovation pace.

- General innovation pace of all 10 companies clearly show "innovation boom" after 1990.

- General innovation trend of top 10 tech companies reflects dot-com crisis as well as financial crisis 2007-2008 with significant down in patent registrations during crisis time.

- Main innovation categories change over time, but "Active solid state devices" is the leading patent category across all periods.
Also the "evolution" of data related leading patent categories is the following: starting from data/information storage before 90th, then data processing: database and file management (1990-2020) to Recording, communication, or information retrieval equipment and image processing (from 2010 till present).

- Innovation direction of tech companies reflects the specific of their businesses:
    
    - Apple, Tencent focus on Recording, communication, or information retrieval equipment
    - Facebook apart from data processing contribute to information security
    - Google is also focued on data processing and contribute to image analysis
    - Samsung, IBM and Intel mostly innovate in hardware
    - Hon Hai Precision innovate in electrical system and devices and in energy transformation equipment.
    - Microsoft and Oracle - digital processing systems and data processing.






